Bayesian Hierarchical Modeling: An Introduction and Reassessment

### 2022-06-29 manuscript.Rmd
Small fixes

fitrstanhiertable32 <- add_column(fitrstanhiertable22, Parameters = rownamesfitrstanhiertable2, .before = "Mean")
apa_table(fitrstanhiertable32, row.names = FALSE, caption = "Posterior Variance, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the Variance Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
```
Thanks to the hierarchical structure of the model, we can investigate whether the general effects hold for all the individuals by inspecting the individual variation and the individual parameter estimates. The individual variation is represented by the variance or standard deviation parameters in the model. The summaries of the posterior distributions of the variance parameters as estimated by `rstan` are shown in Table \@ref(tab:inddevtable). The estimates for the standard deviations by `brms` are smaller compared to `rstan` and can be found in Online Supplement D.

@@

### Model Diagnostics 
Before interpreting the results, we have to check if the posterior distributions of the parameters have converged. The trace plot of $\mu_{\delta_{6}}$ is shown panel A for `rstan` and B for `brms` of Figure \@ref(fig:traceplotlog). Trace plots for all other parameters can be found in the Online Supplement. All plots look like a hairy caterpillar. Therefore, they indicate that the posterior distributions of the parameters have converged. $\hat{R}$ values shown in Figure \@ref(fig:traceplotlog)C for all parameters are all close to 1 and most of the number of effective samples shown in Figure \@ref(fig:traceplotlog)D exceed the total number of iterations. The number of effective samples for the general effects are comparatively low but still sufficient. In combination with the other diagnostics, we conclude that the posterior distributions of the parameters have converged and we proceed with the interpretation of the results.

@@

#### Individual effects 
Next, we inspect the individual deviations from the general effects. Table \@ref(tab:inddevtablelog) presents the estimates of the variance parameters by `rstan`. Results for brms can be found in the Online Supplement. 
```{r inddevtablelog, results = "asis", cache = TRUE}
# example - with SE 
# rstanfithier8log <- readRDS(file = "R objects/hier_logmodelc_adj.rds")
fitrstanhiertable2log <- summary(rstanfithier8log, pars = c("g", "g2", "g3", "g4", "g5", "g6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2log2 <- as.data.frame(fitrstanhiertable2log)
rownamesfitrstanhiertable2log2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")
colnames(fitrstanhiertable2log2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

@@

fitrstanhiertable32log <- add_column(fitrstanhiertable2log2, Parameters = rownamesfitrstanhiertable2log2, .before = "Mean")
apa_table(fitrstanhiertable32log, row.names = FALSE, caption = "Posterior Variance, Lower and Upper Bound of the 95\\% Credible Interval, the Number of Effective Samples, and the $\\hat{R}$ of the Variance Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
```
The estimated parameter values for the variance terms are very small, but different from zero. This indicates that individuals may vary slightly with respect to the general effect parameters. The estimates for the standard deviations by `brms` are similar compared to the estimates by `rstan`, but slightly smaller. The exact estimates by `brms` can be found in Online Supplement F.
Figure \@ref(fig:randomparplotlog) shows the individual estimates for the digit and side effects. The pink lines indicate that the 95% credible interval of the individual effect contains zero. The estimated general effect is represented by the dashed line. For $\beta$, $\delta_{7}$, and $\delta_{3}$, most 95% credible intervals contain zero. For $\delta_{6}$ and $\delta_{4}$, most credible intervals are entirely larger than zero, indicating an increase in RT for most participants.

>>>>

fitrstanhiertable32 <- add_column(fitrstanhiertable22, Parameters = rownamesfitrstanhiertable2, .before = "Mean")
apa_table(fitrstanhiertable32, row.names = FALSE, caption = "Posterior Variance, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the Variance Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
```
Thanks to the hierarchical structure of the model, we can investigate whether the general effects hold for all the individuals by inspecting the individual variation and the individual parameter estimates. The individual variation is represented by the variance or standard deviation parameters in the model. The summaries of the posterior distributions of the variance parameters as estimated by `rstan` are shown in Table \@ref(tab:inddevtable). The estimates for the standard deviations by `brms` are smaller compared to `rstan` and can be found in Online Supplement D and H.

@@

### Model Diagnostics 
Before interpreting the results, we have to check if the posterior distributions of the parameters have converged. The trace plot of $\mu_{\delta_{6}}$ is shown panel A for `rstan` and B for `brms` of Figure \@ref(fig:traceplotlog). Trace plots for all other parameters can be found in Online Supplement G. All plots look like a hairy caterpillar. Therefore, they indicate that the posterior distributions of the parameters have converged. $\hat{R}$ values shown in Figure \@ref(fig:traceplotlog)C for all parameters are all close to 1 and most of the number of effective samples shown in Figure \@ref(fig:traceplotlog)D exceed the total number of iterations. The number of effective samples for the general effects are comparatively low but still sufficient. In combination with the other diagnostics, we conclude that the posterior distributions of the parameters have converged and we proceed with the interpretation of the results.

@@

#### Individual effects 
Next, we inspect the individual deviations from the general effects. Table \@ref(tab:inddevtablelog) presents the estimates of the variance parameters by `rstan`. Results for brms can be found in Online Supplement H. 
```{r inddevtablelog, results = "asis", cache = TRUE}
# example - with SE 
# rstanfithier8log <- readRDS(file = "R objects/hier_logmodelc_adj.rds")
fitrstanhiertable2log <- summary(rstanfithier8log, pars = c("g", "g2", "g3", "g4", "g5", "g6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2log2 <- as.data.frame(fitrstanhiertable2log)
rownamesfitrstanhiertable2log2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")
colnames(fitrstanhiertable2log2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

@@

fitrstanhiertable32log <- add_column(fitrstanhiertable2log2, Parameters = rownamesfitrstanhiertable2log2, .before = "Mean")
apa_table(fitrstanhiertable32log, row.names = FALSE, caption = "Posterior Variance, Lower and Upper Bound of the 95\\% Credible Interval, the Number of Effective Samples, and the $\\hat{R}$ of the Variance Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
```
The estimated parameter values for the variance terms are very small, but different from zero. This indicates that individuals may vary slightly with respect to the general effect parameters. The estimates for the standard deviations by `brms` are similar compared to the estimates by `rstan`, but slightly smaller. The exact estimates by `brms` can be found in Online Supplement F and H.
Figure \@ref(fig:randomparplotlog) shows the individual estimates for the digit and side effects. The pink lines indicate that the 95% credible interval of the individual effect contains zero. The estimated general effect is represented by the dashed line. For $\beta$, $\delta_{7}$, and $\delta_{3}$, most 95% credible intervals contain zero. For $\delta_{6}$ and $\delta_{4}$, most credible intervals are entirely larger than zero, indicating an increase in RT for most participants.

### 2022-06-29 manuscript.tex
Small fixes

Thanks to the hierarchical structure of the model, we can investigate whether the general effects hold for all the individuals by inspecting the individual variation and the individual parameter estimates. The individual variation is represented by the variance or standard deviation parameters in the model. The summaries of the posterior distributions of the variance parameters as estimated by \texttt{rstan} are shown in Table \ref{tab:inddevtable}. The estimates for the standard deviations by \texttt{brms} are smaller compared to \texttt{rstan} and can be found in Online Supplement D.

@@

Before interpreting the results, we have to check if the posterior distributions of the parameters have converged. The trace plot of \(\mu_{\delta_{6}}\) is shown panel A for \texttt{rstan} and B for \texttt{brms} of Figure \ref{fig:traceplotlog}. Trace plots for all other parameters can be found in the Online Supplement. All plots look like a hairy caterpillar. Therefore, they indicate that the posterior distributions of the parameters have converged. \(\hat{R}\) values shown in Figure \ref{fig:traceplotlog}C for all parameters are all close to 1 and most of the number of effective samples shown in Figure \ref{fig:traceplotlog}D exceed the total number of iterations. The number of effective samples for the general effects are comparatively low but still sufficient. In combination with the other diagnostics, we conclude that the posterior distributions of the parameters have converged and we proceed with the interpretation of the results.

@@

Next, we inspect the individual deviations from the general effects. Table \ref{tab:inddevtablelog} presents the estimates of the variance parameters by \texttt{rstan}. Results for brms can be found in the Online Supplement.

@@

The estimated parameter values for the variance terms are very small, but different from zero. This indicates that individuals may vary slightly with respect to the general effect parameters. The estimates for the standard deviations by \texttt{brms} are similar compared to the estimates by \texttt{rstan}, but slightly smaller. The exact estimates by \texttt{brms} can be found in Online Supplement F.
Figure \ref{fig:randomparplotlog} shows the individual estimates for the digit and side effects. The pink lines indicate that the 95\% credible interval of the individual effect contains zero. The estimated general effect is represented by the dashed line. For \(\beta\), \(\delta_{7}\), and \(\delta_{3}\), most 95\% credible intervals contain zero. For \(\delta_{6}\) and \(\delta_{4}\), most credible intervals are entirely larger than zero, indicating an increase in RT for most participants.

>>>>

Thanks to the hierarchical structure of the model, we can investigate whether the general effects hold for all the individuals by inspecting the individual variation and the individual parameter estimates. The individual variation is represented by the variance or standard deviation parameters in the model. The summaries of the posterior distributions of the variance parameters as estimated by \texttt{rstan} are shown in Table \ref{tab:inddevtable}. The estimates for the standard deviations by \texttt{brms} are smaller compared to \texttt{rstan} and can be found in Online Supplement D and H.

@@

Before interpreting the results, we have to check if the posterior distributions of the parameters have converged. The trace plot of \(\mu_{\delta_{6}}\) is shown panel A for \texttt{rstan} and B for \texttt{brms} of Figure \ref{fig:traceplotlog}. Trace plots for all other parameters can be found in Online Supplement G. All plots look like a hairy caterpillar. Therefore, they indicate that the posterior distributions of the parameters have converged. \(\hat{R}\) values shown in Figure \ref{fig:traceplotlog}C for all parameters are all close to 1 and most of the number of effective samples shown in Figure \ref{fig:traceplotlog}D exceed the total number of iterations. The number of effective samples for the general effects are comparatively low but still sufficient. In combination with the other diagnostics, we conclude that the posterior distributions of the parameters have converged and we proceed with the interpretation of the results.

@@

Next, we inspect the individual deviations from the general effects. Table \ref{tab:inddevtablelog} presents the estimates of the variance parameters by \texttt{rstan}. Results for brms can be found in Online Supplement H.

@@

The estimated parameter values for the variance terms are very small, but different from zero. This indicates that individuals may vary slightly with respect to the general effect parameters. The estimates for the standard deviations by \texttt{brms} are similar compared to the estimates by \texttt{rstan}, but slightly smaller. The exact estimates by \texttt{brms} can be found in Online Supplement F and H.
Figure \ref{fig:randomparplotlog} shows the individual estimates for the digit and side effects. The pink lines indicate that the 95\% credible interval of the individual effect contains zero. The estimated general effect is represented by the dashed line. For \(\beta\), \(\delta_{7}\), and \(\delta_{3}\), most 95\% credible intervals contain zero. For \(\delta_{6}\) and \(\delta_{4}\), most credible intervals are entirely larger than zero, indicating an increase in RT for most participants.
